### 2022-04-01 manuscript/thesis.tex
further revisions of section 3

\section{Submission format}
For a collaborative forecast project to run smoothly, forecast submissions from 
all forecasters 
should follow the same format. For both the CDC flu competition and
the COVID-19 Forecast Hub, teams provide a \texttt{.csv} spreadsheet which 
contains the distributional information for one or multiple forecasts. Tables
\ref{table:dstan} and \ref{table:qstan} show what variables are included and a
couple rows to illustrate possible values.
The column variables include \texttt{location}, \texttt{target}, \texttt{type},
\texttt{unit}, \texttt{bin} or \texttt{quantile} and \texttt{value}.
Here \texttt{location} defines the specific county, state, or country of the 
forecast. The \texttt{target} variable defines what is forecasted for example
season onset, deaths, hospitalizations, etc. The \texttt{type} variable defines
the type for the \texttt{value} variable, \texttt{bin} or \texttt{quantile}.
The \texttt{unit} variable defines the time frame of the forecast: week,
two weeks, four weeks, etc. And the \texttt{value} variable is a number that 
either gives probability for a bin or the value for a quantile. Table
\ref{table:dstan} show what a few rows of a forecast 
submission file for the CDC flu competition look like.
For the quantile forecasts of the Forecast Hub, the main changes to Table
\ref{table:dstan} are that the possible values for \texttt{type} 
quantile, and rather than using a \texttt{bin} variable \texttt{quantile} is 
used
with values specified by the Hub. This is seen in Table \ref{table:qstan}

A single submission may include many forecasts each aimed at predicting a 
different combination of \texttt{location}, \texttt{target}, and \texttt{unit}. 
A set of rows which share
the same specific combination of \texttt{location}, \texttt{target}, and 
\texttt{unit} constitute a single forecast. One forecast for the CDC flu
competition may require up to 131 rows whereas in the COVID-19 Forecast Hub one 
forecast may require up to 23 rows.

@@

In the mixture distribution format, an adjustment may be made to the format of 
Tables
\ref{table:dstan} and \ref{table:qstan} where each 
row
represents a component in a mixture distribution.
The variables Bin/Quantile and Value are removed and replaced with Family,
Param1, Param2 and Weight where Family is the distribution family,
Params are the parameters for the component distribution and Weight is the 
weight $w_i$ for that component.

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
 \hline
    location & target & type & unit & family & param1 & param2 & weight
    \\ \hline
    US National & Season Onset & Dist & week & norm & $\mu_n$ & $\sigma_n$ & .\\
    US National & Season Onset & Dist & week & lnorm & $\mu_l$ & $\sigma_l$ &.\\
    ... & ... & ... & ... & ... & ... & ... & .\\
 \hline
 \end{tabular}
 \caption[Mixture distribution forecast submission example]{Example of a 
 submission file
 for a disease outbreak forecast using mixture distribution representation}
 \label{table:mstan}
\end{table}

A forecast center may want to limit the number of components allowed per 
prediction. For reference, a mixture distribution prediction following the 
format in table \ref{table:dstan} with 18 
components would require $18 \times 8 = 144$ cells submitted. This is the same 
number of cells
submitted for a Forecast Hub forecast with 23 quantiles and a point prediction 
according to the format in table \ref{table:dstan}.

Throughout the remainder of this section, explanations of how to work with 
mixture distributions along with \texttt{R} code which will be given to 
demonstrate constructing a mixture distribution from a forecast submission, 
scoring the forecast and building an ensemble forecast.

@@

From a submitted forecast, a prediction may be found by Location, Target and 
Unit (or whatever other indication variables are used by collaborators) and 
evaluated. We found
the \texttt{distr} package \cite{camphausen2007distr}
useful for constructing a mixture distribution given
information like that from table \ref{table:dstan}.

The function 
\texttt{UnivarMixingDistribution} takes as arguments a list of distributions
and a vector of weights for each distribution and a object of class 
\texttt{AbscontDistribution} is returned. The object is in essence a mixture
distribution of the form of (\ref{eq:dmd}). Functions for density, 
distribution
and quantile functions as well as for random sampling may then be used on the
mixture distribution object for evaluation. Using those functions and the 
created object, LogS and CRPS are easily calculable.

We wrote a function \texttt{MakeDist} (see Appendix)
which takes on a data frame with variables
Family, Param1, Param2, Param3 and Weight and returns a mixture distribution 
of component distributions specified in the rows.

Beginning with two forecast submissions of the form of table \ref{table:dstan}, 
say we select from each a prediction of an event for the same location, target
and unit. The following two data frames \texttt{preddf1} and \texttt{preddf2}
are the resulting predictions.

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|}
 \hline
    family & param1 & param2 & param3 & weight
    \\ \hline
    Lnorm & 2 & 1 & NA & 0.3  \\
    Norm & 2.1 & 1 & NA & 0.7 \\
 \hline
 \end{tabular}
 \caption[Toy prediction 1]{Toy prediction 1}
 \label{tab:preddf1}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|}
 \hline
    family & param1 & param2 & param3 & weight
    \\ \hline
    Norm & 1.5 & 1 & NA & 0.4  \\
    Norm & 4 & 2 & NA & 0.6 \\
 \hline
 \end{tabular}
 \caption[Toy prediction 2]{Toy prediction 2}
 \label{tab:preddf2}
\end{table}

The code here shows these two data frames and how the \texttt{MakeDist} 
function is used to create the distributions in \texttt{R}. Once the 
distributions are created as \texttt{AbscontDistribution} classes, then 
functions for evaluating a pdf and a CDF for each are created.

@@

\specialchapt{APPENDIX}

>>>>

\section{Submission format}
For a collaborative forecast project to run smoothly, forecast submissions from 
all forecasters 
should follow the same format. For both the CDC flu competition and
the COVID-19 Forecast Hub, teams provide a \texttt{.csv} spreadsheet which 
contains the distributional information for one or multiple forecasts. Tables
\ref{table:dstan} and \ref{table:qstan} show what variables are included and a
couple rows to illustrate possible values.
The column variables include \texttt{location}, \texttt{target}, \texttt{type},
\texttt{unit}, \texttt{bin} or \texttt{quantile}, and \texttt{value}.
Here \texttt{location} defines the specific county, state, or country of the 
forecast. The \texttt{target} variable defines what is forecasted with levels
season onset, deaths, hospitalizations, etc. The \texttt{type} variable defines
the type for the \texttt{value} variable with levels of \texttt{point}, 
bin, or quantile.
The \texttt{unit} variable defines the time frame of the forecast with levels of
week,
two weeks, four weeks, etc. The variables \texttt{bin} and \texttt{quantile}
give a specific bin or a specific quantile. And the \texttt{value} variable is a 
number that 
either gives the probability associated with a bin or the value associated 
with a quantile.

A single submission may include many forecasts aimed at forecasting a 
different combinations of \texttt{location}, \texttt{target}, and \texttt{unit}. 
A set of rows which share
the same specific combination of \texttt{location}, \texttt{target}, and 
\texttt{unit} constitute a single forecast. One forecast for the CDC flu
competition may require up to 131 rows whereas in the COVID-19 Forecast Hub one 
forecast may require up to 23 rows.

@@

Table \ref{table:mstan} illustrates adjustments made to the submission formats
from 
Tables \ref{table:dstan} and \ref{table:qstan} which make a useable submission 
format for mixture distribution forecasts. In such a format, each row represents
a component used in a mixture distribution.
The variables \texttt{bin} or \texttt{quantile} and \texttt{value} are removed 
and replaced with \texttt{family},
\texttt{param1}, \texttt{param2}, and \texttt{weight} where \texttt{family} is 
the distribution family of the component,
\texttt{param1} and \texttt{param2} are the parameters for the component 
distribution, and \texttt{weight} is the 
weight $w_i$ for that component.

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
 \hline
    location & target & type & unit & family & param1 & param2 & weight
    \\ \hline
    US National & Season Onset & Dist & week & norm & $a_n$ & $b_n$ & $w_1$\\
    US National & Season Onset & Dist & week & lnorm & $a_l$ & $b_l$ & $w_2$\\
    ... & ... & ... & ... & ... & ... & ... & ...\\
 \hline
 \end{tabular}
 \begin{minipage}{11cm}
\captionsetup{font=scriptsize}
 \caption[Mixture distribution forecast submission example]{This is an example 
 of a 
 submission file
 for a disease outbreak forecast using a mixture distribution representation.
 Each row represents a component of a mixture distribution. The variables
 location, target, and unit specify what is being forecasted. The variable type
 specifies that the row represents a parametric distribution. The variables
 family, param1, and param2 specify the exact distribution. And the variable 
 weight specifies the weight $w_i$ that the component distribution has in the
 mixture distribution. Here two components are shown with distributions
 Normal$(a_n, b_n)$, Lognormal$(a_l, b_l)$ and weights $w_1$ and $w_2$.}
 \label{table:mstan}
 \end{minipage}
\end{table}

For reasons of storage and computation, a forecast project may have a limit to 
the number of components allowed per forecast.
For reference, a mixture distribution forecast following the 
format in table \ref{table:dstan} with 17
components would require $17 \times 8 = 136$ cells submitted per forecast. 
A submission to the COVID-19 Forecast Hub forecast 23 quantiles
according to the format in table \ref{table:dstan} requires
$23 \times 6 = 138$ cells. Thus if the COVID-19 Forecast Hub were to change the
forecast representation from quantile forecasts to mixture forecasts but 
continue allowing the same amount of data per forecast, 
a mixture distribution with 17 components could be used in one forecast. That
many components could allow for a large range of distribution shapes and 
flexible forecasts.

Throughout the remainder of this section, explanations of how to work with 
mixture distributions submitted according to Table \ref{table:mstan} is 
given. Also given is
\texttt{R} code which 
demonstrates constructing a mixture distribution from a forecast submission, 
scoring the forecast, and building an ensemble from two separate submissions.

@@

A single \texttt{.csv} submission file of the format in Table \ref{table:mstan} 
may contain multiple forecasts forecasting different combinations of location,
target, and unit. Selecting only rows which share a specific combination of 
location, target, and unit will produce a table representing a single forecast.
That table may look like Tables \ref{tab:preddf1} and 
\ref{tab:preddf1}. If the table is saved as a standard \texttt{data.frame} in 
\texttt{R}, then tools based on the \texttt{distr} package 
\cite{camphausen2007distr} may be used for evaluating a mixture distribution 
with the components in the table.

The \texttt{distr} package contains a function 
\texttt{UnivarMixingDistribution()} takes as arguments a list of distributions
and a vector of weights for each distribution and a object of class 
\texttt{AbscontDistribution} is returned. An \texttt{AbscontDistribution} class
is a mother class which defines a random number generator, pdf,
CDF, and quantile function for continuous distributions from common
families contained in the \texttt{distr} and for mixture distributions with 
component distributions from those families.
We wrote a function \texttt{MakeDist()} (see APPENDIX)
which takes on a \texttt{data.frame} with variables
family, param1, param2, param3, and weight and where each row represents a
component distribution in a mixture distribution. The function 
\texttt{MakeDist()} calls on the \texttt{UnivarMixingDistribution()} function
and returns a mixture distribution object of class \texttt{AbscontDistribution}.

If a forecast such as in Table \ref{tab:preddf1} is taken as an argument in 
\texttt{MakeDist()}, the resulting mixture distribution may then be evaluated 
with functions for the pdf, CDF, quantile function, and random samples from the
mixture distribution may be drawn. Thus the distribution may be scored using the
LogS or CRPS.

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|}
 \hline
    family & param1 & param2 & param3 & weight
    \\ \hline
    Lnorm & 2 & 1 & NA & 0.3  \\
    Norm & 2.1 & 1 & NA & 0.7 \\
 \hline
 \end{tabular}
 \caption[Illustrative forecast 1]{Toy prediction 1}
 \label{tab:preddf1}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{|c|c|c|c|c|}
 \hline
    family & param1 & param2 & param3 & weight
    \\ \hline
    Norm & 1.5 & 1 & NA & 0.4  \\
    Norm & 4 & 2 & NA & 0.6 \\
 \hline
 \end{tabular}
 \caption[Illustrative forecast 2]{Toy prediction 2}
 \label{tab:preddf2}
\end{table}

Here we include code to illustrate the process of constructing and scoring two 
separate forecasts. We suppose that Table \ref{tab:preddf1} represents a 
submitted forecast from one forecaster and Table \ref{tab:preddf2} represents
a forecast of the same event from a second forecaster.
Note the additional \texttt{param3} variable
in Tables \ref{tab:preddf1} and \ref{tab:preddf1}. This variable is included
in the table because of the functionality of the \texttt{MakeDist()} function
which allows for component distributions of three parameters.
The code here shows these two forecasts as \texttt{data.frame}s and how the 
\texttt{MakeDist()} 
function is used to create the distributions in \texttt{R}. Once the 
distributions are created as \texttt{AbscontDistribution} objects, then 
functions for evaluating a pdf and a CDF for each are created.

@@

\specialchapt{APPENDIX}
\label{section:append}
